# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}
for (i in 1:50){
  print(i)
}
# any i between 1 to 50 will be printed. For example, for i=1 it will do the print function; for i=2 it will do the print functiion; The loop does not include i=51, so the print function stop.
```



### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
s=0
for (i in 0:1000){
  s=s+i
}
print(s)
# for any i between 1 to 1000, it will add i to the s. for example, for i=0 s=0; for i =1, s=0+1=1; fori=2, s=1+2=3; for i=3,s=3+3=6; it will be done when i=1000. The loop does not include i=1001, so the summation is done for i=1000.
```



B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
s=0
for (i in seq(0,1000,by=2)){
  s=s+i
  }
print(s)
# for any even integers i  between 1 to 1000, it will add i to the s. for example, for i=0 s=0; for i =2, s=0+2=2; fori=4, s=4+2=6; for i=6,s=6+6=12; it will be done when i=1000. The loop does not include i=1002, so the summation is done for i=1000.

```




C. Now, repeat A and B WITHOUT using a loop.
```{r}
print(sum(c(0:1000)))
# get a sequence from 0 to 1000, and use sum to add all the numbers in the sequence, and then print the number#
print(sum(seq(0,1000,by=2)))
# get a sequence of even integers from 0 to 1000, and use sum to add all the even numbers in the sequence, and then print the number#
```


### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
for(i in 1:6){ # it has six rows in the dataframe
  for(j in 1:6){ # it has six column in the dataframe
    if(survey[i,j]<1 |survey[i,j]>5){ # if the number in the cell is smaller than 1 or the number in the cell is larger than 5
    survey[i,j]=NA # it will be replaced by NA, if the number in the cell is between 1 to 5 (include 1 and 5), it will not be replaced.
    } 
 }
}

```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.

```{r}
library(dplyr)
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2))

for(i in 1:6){# it has six rows in the dataframe
  for(j in 1:6){# it has six column in the dataframe
    if(survey[i,j]<1 |survey[i,j]>5){# if the number in the cell is smaller than 1 or the number in the cell is larger than 5
    survey[i,j]=NA# it will be replaced by NA, if the number in the cell is between 1 to 5 (include 1 and 5), it will not be replaced.
    } 
    survey$invalid.answers <- apply(is.na(survey), 1, sum) # after all the numbers in the cell between 1 and 5 (include 1 and 5), create a new column in the data frame called "invalid.answers" and the numbers in each cell means how many NA in its row.

  }
  
}


```






