---
title: "Simulations Pre-Class Project"
output:
  html_document: default
  pdf_document: default
date: "Due March 13, 2017 at 5:00pm"
---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
```{r}

gambler_ruin=function(bankroll,bet,k0) # creat a function, with input bankroll, bet and number of hands
{ 
  k=0 # at the begining the count of hand is 0
  while (bankroll>=bet & k<k0){# if bankroll larger or equal to bet, which means the gambler has money to gamble, and the count of hand is smaller than the max hand he could have, then do the loop, or stop
   
    if(rbinom(1, 1, 0.5) == 1) # generate a number which 50% could be 1, 50% could be 0. If it is 1, then the gambler wins and get bet back and get extra money which equals to bet
    {
      bankroll = bankroll+bet
      
    }else{
      bankroll = bankroll-bet # if the number is not 1, which means the gambler loses, then the bankroll will lose the bet.
      }
    k=k+1 # the count of hand plus 1, when one hand done.
    
  
  }
  return(c(bankroll,k)) # return how much bankroll the gambler gets and how many counts of hand he does. if the bankroll equals 0, which means he go bust and we want to know after how many hands, he goes bust. if the bankroll does not equal to 0, which means he does not go bust, and we want to know how much money he gets finally after the maxi hands.
}

a100=replicate(5000, gambler_ruin(1000,100,100)) # when the gambler plays 100 hands (maximum), the money and the hands he could did. then we simulate it 5000 times
b100=sum(a100[1,]==0)/5000 # when the bankroll equals zero, which means the gambler goes bust, we count how many times this happened in the simulation and divide it by 5000 to get the possiblity
b100

```

b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
```{r}
a500=replicate(5000, gambler_ruin(1000,100,500))# when the gambler plays 500 hands (maximum), the money and the hands he could did. then we simulate it 5000 times
b500=sum(a500[1,]==0)/5000# when the bankroll equals zero, which means the gambler goes bust, we count how many times this happened in the simulation and divide it by 5000 to get the possiblity
b500
```    
    c. the mean time you go bust, given that you go bust within the first 5000 hands.
```{r}
c500=numeric() # creat a vector 
c100=numeric() # creat a vector
for (i in 1:5000){ #5000 simulations
  if (a500[1,i]==0){ # if the bankroll equals to 0, which means he goes bust, then add the hands, he did to a the vector
   c500=c(c500, a500[2,i])
 }
}
for (i in 1:5000){# if the bankroll equals to 0, which means he goes bust, then add the hands, he did to a the vector
  if (a100[1,i]==0){
   c100=c(c100, a100[2,i])
 }
}
 
c100_mean=mean(c100) # calucalte the mean of the hands
c500_mean=mean(c500)# calucalte the mean of the hands

c100_mean
c500_mean

```
    
    d. the mean and variance of your bankroll after 100 hands (including busts).
   
```{r}

d_mean=mean(a100[1,]) #the mean of your bankroll after 100 hands
d_var=var(a100[1,])# the variance of your bankroll after 100 hands
d_mean
d_var
```     
 e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
e_mean=mean(a500[1,])#the mean of your bankroll after 500 hands
e_var=var(a500[1,])# the variance of your bankroll after 500 hands
e_mean
e_var
```

Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?



2. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.

```{r}

Markov_Chains=function(money,p0,increase)
{
  k=0 # count =0 at the beginning
  p=p0 # intinal possiblity
  while (money>=100 & k<= 100000){ # the money has be more than wager to gamble, it has to be less than 100000 hands
   
    
    if(rbinom(1, 1, p) == 1) # the probility of winning is p
    {
      money = money+100 # if he wins, he get $100 more
      p = p+increase # if he wins ,the possiblity of winning next time increases 1%
      if(p>=1){p=1} # the possiblity could not be more than 1
    }else{
      money = money-100 # if he loses, he will lose the wager $100
      p = p0}
    k=k+1 # count plus 1
  
  }
  return(money)
}

a=replicate(100, Markov_Chains(500000,0.48,0.01)) # repeat 100 times
a
mean(a) # mean of the final result
```






    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
It is not a fair game. from the simulation , the gambler's money is always less than the inital value,(less than 500000). Therefore it is not a fair game.
    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
    
    
```{r}

set.seed(1)
fair_probility=function(money,p0,increase){
  expectation=mean(replicate(100, Markov_Chains(money,p0,increase))) # calulate the mean for the initial value, p0=0.48,increas=0.01
  mid=p0 # inital middle possiblity =0.48
  high=p0+0.02 # the upper boundry of the possiblity=0.48
  low=p0-0.02 # the lower boundry of the possiblity=0.46
  while (expectation>money*1.02 | expectation<money*0.98){ # if the mean value(expectaion) is outside the value we expect (0.98*money, 1.02*money), run the loop
    if (expectation > money*1.02) { # if the expecataion is larger than we expect, means, the possiblit is higher than it needs to be, than use the middle value between mid and low to be the new possiblity
      high <- mid
      mid <- (low+high)/2
    } else {# if the expecataion is smaller than we expect, means, the possiblit is lower than it needs to be, than use the middle value between mid and high to be the new possiblity
      low <- mid
      mid <- (low+high)/2
    }
      expectation=mean(replicate(100, Markov_Chains(money,mid,increase))) # run the markov_charins again to get a new expectation (mean of money) and compare it to what we expect to decide if run the loop again
      print(paste("step before final result: ",c(mid,expectation))) # print each step
  }
  return(c(mid,expectation))
}
          
b=fair_probility(500000,0.48,0.01)  
b
```
    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 



 
 
```{r}
fair_increase=function(money,p0,increase){
  expectation1=mean(replicate(100, Markov_Chains(money,p0,increase))) # use the inital value increase =0.01 to calculate a expectation
  mid=increase # inital increase =0.01
  high=1 # the upper boundry of increase=1
  low=0 # the lower boundry of increase=0
  while (expectation1>money*1.02 | expectation1<money*0.98){ # if the expectation is outside the range, then run the loop
     if (expectation1 > money*1.02){  # if the expectation is larger than 2% ,which means the increase is higher than it needs to be, then we need a lower increase. Therefore we find the average of mid and low.
      high <- mid
      mid <- (low+high)/2
    }
  else{# if the expectation is smaller than or equal to 2% ,which means the increase is lower than it needs to be , then we need a higher increase. Therefore we find the average of mid and high.
    low <- mid
    mid <- (low+high)/2
  }
  expectation1=mean(replicate(100, Markov_Chains(money,p0,mid))) # calculate a new expectation
  print(paste("step before final result: ",c(mid,expectation1))) # print each step
  } 
  return(c(mid,expectation1)) 
}
c=fair_increase(500000,0.48,0.01)  
print(paste("fair result is",c))
```
    
    
    
    
    
    
    
